Cost of Living in Toronto

### 2024-09-22 paper/paper.qmd
added introduction

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages('opendatatoronto')
#install.packages("dplyr")
#install.packages('tidyverse')
#install.packages('sf')
#remove.packages('knitr')
install.packages('xfun')
install.packages('knitr')

@@

## Read the 2 cleaned CSV files ##
analysis_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))

analysis_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
```

@@

While this may seem like a perfect place to live for many international talents, residents of Toronto have said this to be far from the truth, with many sturggling financially to meet ends.

# Data
```{r}

analysis_data
#kable(analysis_data)
#knitr::kable(head(analysis_data[, 1:4]), format = "pipe")
analysis_data %>% head() %>% knitr::kable()
```
## Data Source

@@

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages('opendatatoronto')
#install.packages("dplyr")
#install.packages('tidyverse')
#install.packages('sf')
#remove.packages('knitr')
#install.packages('xfun')
#install.packages('knitr')

@@

## Read the 2 cleaned CSV files ##
cleaned_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))

cleaned_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
```

@@

While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many sturggling financially to meet ends, especially for those that come from a low income background. For example, according to the Daily Bread Food Bank, one in 10 people in Toronto relied on food banks in 2023, a huge increase from one in 20 from 2022 (CITATION HERE).

According to the City of Toronto, the threshold for LIM_AT (threshold to be considered low income) was $26,503 for 1 person household, $37,480 for two person household and $53,005 for a four person household (CITATION, CITY OF TORONTO). 13.2% of Toronto's population fell under this category of low-income. Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at 8.09% (CITATION HERE), highlighting an issue of income discrepancy.

This paper aims to analyze if low income families living in Toronto are financially able to survive. The analysis focuses on two different metrics; income (which includes income from employment, government benefits and transfers such as GST/HST credit, Canada Pension Plan and Canada Worker Benefits) and expense (which includes rent, childcare, food and transportation costs). The difference between income and expense can assess whether or not low-income families are able to survive, and can guide future policymakers on what areas of expenses the low-income families can get additional support.

The paper is organized as follows: Section 2 (Data) of this paper introduces the raw dataset of interest and some of the initial observations. Section 3 (Results) highlights some of the key insights that were derived from the data analysis. Section 4 (Discussion) discusses the observations from Section 3 and draws a conclusion from the findings. Finally, Section 5 (Conclusion) draws a brief summary of the key findings.

# Data
```{r}
#| label: low-income-families
#| fig-cap: "Table 1: Sample of the Income/Expenses of Low Income Households in Toronto"
#| echo: false

# PREVIEW OF THE DATA
cleaned_data[, 1:5] %>% head() %>% knitr::kable()
```
## Data Source

@@

https://www.toronto.ca/wp-content/uploads/2022/07/9877-City-Planning-2021-Census-Backgrounder-Families-Hhlds-Marital-Status-Income.pdf

https://www.dailybread.ca/wp-content/uploads/2023/11/DB-WhosHungryReport-2023-Digital.pdf
